---
title: "Counterfactual thought without counterfactual langauge"
author: "David Rose, Siying Zhang, Hyowon Gweon, Tobias Gerstenberg"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  bookdown::html_document2:
    toc: true
    toc_depth: 6
    toc_float: true
    theme: cosmo
    highlight: tango
---

# Load packages

```{r setup, include = F, message = F, warning = F}

library("xtable")         # for saving tables
library("emmeans")        # for comparing means
library("knitr")          # for knitting  
library("brms")           # for Bayesian data analysis
library("tinytable")      # for tidy tables
library("broom.mixed")    # for tidy model results
library("boot")           # for inv.logit() function
library("DT")             # for datatable() function 
library("RColorBrewer")   # for color palettes
library("tidyverse")      # for everything else 


```

# Settings
```{r}


knitr::opts_chunk$set(echo = T, warning = F, message = F)

# Define a custom color palette
children_palette = brewer.pal(n = 9, name = "Greens")
adults_palette = "grey"
colors = c(children_palette, adults_palette)

# set theme 
theme_set(theme_classic())

# suppress grouping warning 
options(dplyr.summarise.inform = F)

```


# Helper functions

## Demographics for children
```{r}
print_demographics = function(data) {
  # gender
  data %>%
    group_by(gender) %>%
    summarise(n = n_distinct(participant)) %>%
    print()
  
  # age
  print('age:')
  mean(data$age, na.rm = T) %>% print()
  sd(data$age, na.rm = T) %>% print()
  
  # language spoken
 data %>%
    mutate(en_occurrence = ifelse(str_detect(lan_spoken, "en"), "en", "non-en")) %>%
    group_by(en_occurrence) %>%
    summarise(n = n_distinct(participant)) %>%
    print()

}

```


## Demographics for adults
```{r}
print_demographics_adults = function(data) {
  # gender
  data %>%
    group_by(gender) %>%
    summarise(n = n()) %>%
    print()
  
  # age
  print('age:')
  mean(data$age, na.rm = T) %>% print()
  sd(data$age, na.rm = T) %>% print()

# race
  data %>%
    group_by(race) %>%
    summarise(n = n()) %>%
    print()
  
}

```

# EXPERIMENT 1 
## CHILDREN
### DATA
#### Read in Data

```{r message=FALSE, warning=FALSE}

df.exp1_children = read_csv("../../data/experiment1/children/exp1_children.csv") %>%
  filter(row_number() >= 0 & row_number() <= 120) %>% 
  select(age, gender, lan_spoken, trial_1, trial_1_score, trial_2, trial_2_score, trial_3, trial_3_score, trial_4, trial_4_score, control_trial_score, total_score) %>%
  mutate(age_group = cut(age, breaks = c(3, 4, 5, 6, 7.1),
                         labels = c("3", "4", "5", "6"),
                         right = TRUE)) %>%
  # na.omit() %>%
  mutate(participant = 1:n(),
         total_score = as.numeric(total_score),
        across("trial_1"|"trial_2"|"trial_3"|"trial_4", ~case_when(
    grepl("(andy|emily)_correct_(left|right)_(andy|emily)", .) ~ "egg | basketball",
    grepl("(taylor|zoey)_correct_(left|right)_(taylor|zoey)", .) ~ "phone | rubberducky",
    grepl("(sara|leo)_correct_(left|right)_(sara|leo)", .) ~ "vase | toybear",
    grepl("(olivia|mason)_correct_(left|right)_(olivia|mason)", .) ~ "icecream | sunscreen",
    TRUE ~ .
  ))) %>%
  mutate(across(starts_with("trial"), as.character)) %>%
  pivot_longer(cols = c("trial_1", "trial_2", "trial_3", "trial_4"),
               names_to = "trial_order",
               values_to = "trial_name") %>%
  group_by(participant) %>%
  mutate(trial_score = case_when(
    trial_order == "trial_1" ~ trial_1_score,
    trial_order == "trial_2" ~ trial_2_score,
    trial_order == "trial_3" ~ trial_3_score,
    trial_order == "trial_4" ~ trial_4_score,
  )) %>%
  mutate(trial_score = as.numeric(trial_score)) %>%
  mutate(trial_order = ifelse(trial_order == "trial_1", "trial 1",
                               ifelse(trial_order == "trial_2", "trial 2",
                                      ifelse(trial_order == "trial_3", "trial 3",
                                             ifelse(trial_order == "trial_4", "trial 4", trial_order))))) %>%
  select(participant, gender, lan_spoken, age, age_group, trial_order, trial_name, trial_score, total_score, control_trial_score)

```

#### Demographics
```{r}

print_demographics(df.exp1_children)

```

### STATS
##### Bayesian Regression Models
###### Hypothesis 1 
```{r}

# Overall children will thank the individual who prevented the worse outcome from occurring at rates that exceed chance

fit.brm1 = brm(formula = trial_score ~ 1 + (1 | participant) + (1 | trial_name),
     data = df.exp1_children,
     family = "bernoulli", 
     cores = 4,
     file = "cache/brm1") 

fit.brm1 %>% 
  tidy(effects = c("fixed"),
       conf.int = TRUE,
       conf.level = 0.95,
       conf.method = "HPDinterval") %>% 
  mutate(across(.cols = c(estimate, conf.low, conf.high),
                .fns = ~ inv.logit(.))) %>% 
  select(estimate, contains("conf")) 
  

```

###### Hypothesis 2
```{r}

# The tendency to thank the individual who prevented the worse potential outcome from occurring will increase with age

fit.brm2 = brm(formula = trial_score ~ 1 + age + (1 | participant) + (1 | trial_name),
     data = df.exp1_children,
     family = "bernoulli", 
     cores = 4,
     file = "cache/brm2") 

fit.brm2 %>% 
  tidy(effects = c("fixed"),
       conf.int = TRUE,
       conf.level = 0.95,
       conf.method = "HPDinterval") %>% 
  filter(term == "age") %>%
  select(estimate, contains("conf"))
 
```

## ADULTS
### DATA
#### Read in data


```{r message=FALSE, warning=FALSE}

df.exp1_adults = read_csv("../../data/experiment1/adults/exp1_adults.csv")

```

#### Wrangle 

```{r}

df.exp1_adults = df.exp1_adults %>% 
  mutate(participant = 1:30) %>% 
  relocate(participant) %>% 
  select(-workerid:-post_test_trial, -error) %>% 
   pivot_longer(cols = c("video_1", "video_2", "video_3", "video_4"
   ),
   names_to = "video_version",
   values_to = "video"
  ) %>% 
  pivot_longer(cols = c("video_1_correct_answer", "video_2_correct_answer", "video_3_correct_answer", "video_4_correct_answer"),
               names_to = "video_version_correct_answer",
               values_to = "correct_answer") %>% 
  pivot_longer(cols = c("video_1_response", "video_2_response", "video_3_response", "video_4_response"),
                names_to = "video_version_response",
   values_to = "response") %>% 
  mutate(video_version_correct_answer = str_remove_all(video_version_correct_answer, "_correct_answer"),
         video_version_response = str_remove_all(video_version_response, "_response")) %>% 
  filter(video_version == video_version_correct_answer & video_version_correct_answer == video_version_response) %>% 
  select(participant, scenario = video, correct_answer, response) %>% 
  mutate(rating = if_else(correct_answer == response, 1, 0))

```

#### Demogrpahics

```{r}
# read in data
df.exp1_adults_demographics = read_csv("../../data/experiment1/adults/exp1_adults_demographics.csv")

# print demographics
print_demographics_adults(df.exp1_adults_demographics)


```

### STATS
#### Bayesian Regression Models
##### Hypothesis 1 
```{r}

# Overall adults will thank the individual who prevented the worse outcome from occurring at rates that exceed chance

# fit.brm1.adults = brm(formula = response ~ 1 + (1 | participant) + (1 | scenario),
#      data = df.exp1_adults,
#      family = "bernoulli",
#      cores = 4,
#      file = "cache/brm1.adults") 
# 
# fit.brm1.adults %>% 
#   tidy(effects = c("fixed"),
#        conf.int = TRUE,
#        conf.level = 0.95,
#        conf.method = "HPDinterval") %>% 
#   mutate(across(.cols = c(estimate, conf.low, conf.high),
#                 .fns = ~ inv.logit(.))) %>% 
#   select(estimate, contains("conf")) 
```


### PLOTS
#### Adults + children
```{r fig.height=4, fig.width=6.5, message=FALSE, warning=FALSE}


selected_indices = seq(from = 1, to = 480, by = 4)  
selected_data = df.exp1_children[selected_indices, ]

# children total counterfactual
df.selected.children.individual = selected_data %>%
    group_by(participant, age_group, age) %>% 
    summarize(pct_correct = mean(total_score) / 4,
              age_mean = mean(age)) %>% 
    mutate(age = as.numeric(age)) %>%
    ungroup() 


# means by age for children
df.children.means = selected_data %>%
  mutate(pct_correct = total_score / 4) %>% 
  group_by(age_group) %>%
  summarize(
    n = n(),
    age_mean = mean(age),
    response = Hmisc::smean.cl.boot(pct_correct)) %>%
    mutate(index = c("mean", "low", "high")) %>%
    ungroup() %>% 
    pivot_wider(names_from = index,
                values_from = response) 



# adult total counterfactual
df.adults.individual = df.exp1_adults %>%
  rename(trial_name = scenario) %>%
  group_by(participant) %>%
  summarize(
    total_score = sum(rating),
    pct_correct = mean(total_score) / 4) %>%
  ungroup() %>%
  mutate(age_group = "adults",
         age_mean = "8",
         age = "8") %>%
  select(-total_score)


# combine total counterfactual data for adults + children
df.combined.individual = rbind(df.selected.children.individual, df.adults.individual)


# adult means
df.adults.means = df.adults.individual %>%
  group_by(age_group) %>%
  summarize(
    n = n(),
    response = Hmisc::smean.cl.boot(pct_correct)) %>%
    mutate(index = c("mean", "low", "high")) %>%
    ungroup() %>% 
    pivot_wider(names_from = index,
                values_from = response) %>%
    mutate(age_mean = as.numeric(8))


# combine means for adults + children
df.combinbed.means = rbind(df.children.means, df.adults.means)

# plot text showing n for adults + children
df.combined.text = df.combinbed.means %>% 
    select(age_group, age_mean, n) %>% 
    mutate(label = n,
           label = ifelse(age_group == "3", str_c("n = ", n), n),
           y = 0.95)

ggplot() + 
  geom_hline(yintercept = seq(0, 1, 0.25),
             linetype = "dotted",
             color = "black",
             alpha = 0.8) +
  geom_point(data = df.selected.children.individual,
             mapping = aes(x = age,
                           y = pct_correct,
                           color = age),
             size = 1.5,
             alpha = 1) +
  geom_pointrange(data = df.combinbed.means %>% filter(age_group != "adults"),
                  mapping = aes(x = age_mean,
                                y = mean,
                                ymin = low,
                                ymax = high,
                                fill = age_mean),
                  shape = 21, 
                  color = "black", 
                  size = .7,
                  show.legend = FALSE,
                  alpha = 0.8) +
  geom_pointrange(data = df.combinbed.means %>% filter(age_group == "adults"),
                  mapping = aes(x = age_mean,
                                y = mean,
                                ymin = low,
                                ymax = high),
                  shape = 23, 
                  fill = "grey",
                  color = "black",
                  size = .7,
                  show.legend = FALSE,
                  alpha = 0.8) +
  geom_text(data = df.combined.text,
            mapping = aes(x = age_mean,
                          y = y,
                          label = label),
            size = 5,
            color = "black",
            vjust = -1.7) +  
  ggtitle("Who should receive the thank-you sticker?") +
  labs(x = "Age (in years)",
       y = "Probability of thanking the person\n who prevented a negative outcome") +
  coord_cartesian(ylim = c(0, 1)) +
  scale_y_continuous(breaks = seq(0, 1, 0.25),
                     labels = str_c(seq(0, 1, 0.25) * 100, "%"),
                     expand = expansion(add = c(0, 0.12))) + 
  scale_x_continuous(breaks = c(seq(3, 7, 1), 8),
                     labels = c(seq(3, 7, 1), "adult"), 
                     expand = expansion(add = c(0.2, 0.5))) +
  scale_fill_gradientn(colors = children_palette) + 
  scale_color_gradientn(colors = children_palette) + 
  theme(
    axis.text.x = element_text(size = 11),
    axis.text.y = element_text(size = 11),
    axis.title = element_text(size = 11),
    plot.title = element_text(size = 12,
                              hjust = 0.5,
                              face = "bold",
                              margin = margin(b = 15)),
    legend.position = "none",
    legend.text = element_text(size = 11),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    strip.text.x = element_text(margin(b = 0, t = 7, l = 2, r = 3)),
    plot.margin = margin(t = 0.5,
                         l = 0.5,
                         r = 0.5,
                         b = 0.5,
                         unit = "cm"))

ggsave("../../figures/experiment1/exp1_results.pdf",  width = 6.5, height = 4)

```

# EXPERIMENT 2 
## CHILDREN
### DATA
#### Read in Data
```{r message=FALSE, warning=FALSE}

df.exp2_children = read_csv("../../data/experiment2/children/exp2_children.csv") %>%
  filter(row_number() >= 0 & row_number() <= 120) %>% 
  select(age, gender, lan_spoken, trial_1, trial_1_score, trial_2, trial_2_score, trial_3, trial_3_score, trial_4, trial_4_score, trial_5, trial_5_score, trial_6, trial_6_score, total_score) %>%
  mutate(age_group = cut(age, breaks = c(3, 4, 5, 6, 7.1),
                         labels = c("3", "4", "5", "6"),
                         right = TRUE)) %>%
  na.omit() %>%
  mutate(participant = 1:n(),
         total_score = as.numeric(total_score),
        across("trial_1"|"trial_2"|"trial_3"|"trial_4"|"trial_5"|"trial_6", ~case_when(
    grepl("lightbulb_laundry_correct_(left|right)_(jessie|caleb)", .) ~ "lightbulb",
    grepl("apple_trashcan_correct_(left|right)_(lily|alex)", .) ~ "apple",
    grepl("basketball_egg_correct_(left|right)_(charlie|noah)", .) ~ "basketball",
    grepl("glass_beanbag_correct_(left|right)_(emily|lucas)", .) ~ "glass",
    grepl("rock_wine_correct_(left|right)_(amber|billy)", .) ~ "rock",
    grepl("bear_water_correct_(left|right)_(vicky|connor)", .) ~ "bear",
    TRUE ~ .
  ))) %>%
  mutate(across(starts_with("trial"), as.character)) %>%
  pivot_longer(cols = c("trial_1", "trial_2", "trial_3", "trial_4", "trial_5", "trial_6"),
               names_to = "trial_order",
               values_to = "trial_name") %>%
  group_by(participant) %>%
  mutate(trial_score = case_when(
    trial_order == "trial_1" ~ trial_1_score,
    trial_order == "trial_2" ~ trial_2_score,
    trial_order == "trial_3" ~ trial_3_score,
    trial_order == "trial_4" ~ trial_4_score,
    trial_order == "trial_5" ~ trial_5_score,
    trial_order == "trial_6" ~ trial_6_score,
  )) %>%
  mutate(trial_score = as.numeric(trial_score)) %>%
  mutate(trial_order = ifelse(trial_order == "trial_1", "trial 1",
                               ifelse(trial_order == "trial_2", "trial 2",
                                      ifelse(trial_order == "trial_3", "trial 3",
                                             ifelse(trial_order == "trial_4", "trial 4",
                                                    ifelse(trial_order == "trial_5", "trial 5",
                                                           ifelse(trial_order == "trial_6", "trial 6", trial_order))))))) %>%
  mutate(side = case_when(
    trial_name %in% c("basketball", "bear", "apple", "rock") ~ "side with things",
    trial_name %in% c("lightbulb", "glass") ~ "side with nothing",
    TRUE ~ "Uncategorized" 
  )) %>%
  mutate(prefer_things = case_when(
    side == "side with things" & trial_score == 1 ~ 1,
    side == "side with nothing" & trial_score == 0 ~ 1,
    TRUE ~ 0
  )) %>%
  select(participant, gender, lan_spoken, age, age_group, trial_order, trial_name, trial_score, side, prefer_things, total_score)

```

#### Demographics
```{r}

print_demographics(df.exp2_children)

```

### STATS
#### Bayesian Regression Models
##### Hypothesis 1 
```{r message=FALSE, warning=FALSE}

# Overall children will thank the individual who prevented the worse outcome from occurring at rates that exceed chance

fit.brm3 = brm(formula = trial_score ~ 1 + (1 | participant) + (1 | trial_name),
     data = df.exp2_children,
     family = "bernoulli", 
     cores = 4,
     file = "cache/brm3") 

fit.brm3 %>% 
  tidy(effects = c("fixed"),
       conf.int = TRUE,
       conf.level = 0.95,
       conf.method = "HPDinterval") %>% 
  mutate(across(.cols = c(estimate, conf.low, conf.high),
                .fns = ~ inv.logit(.))) %>% 
  select(estimate, contains("conf"))

```

##### Hypothesis 2
```{r message=FALSE, warning=FALSE}

# The tendency to thank the individual who prevented the worse potential outcome from occurring will increase with age

fit.brm4 = brm(formula = trial_score ~ 1 + age + (1 | participant) + (1 | trial_name),
     data = df.exp2_children,
     family = "bernoulli", 
     cores = 4,
     file = "cache/brm4") 

fit.brm4 %>% 
  tidy(effects = c("fixed"),
       conf.int = TRUE,
       conf.level = 0.95,
       conf.method = "HPDinterval") %>% 
  filter(term == "age") %>%
  select(estimate, contains("conf")) 

```

## ADULTS
### DATA
#### Read in Data

```{r message=FALSE, warning=FALSE}

# import raw data
df.exp2_adults = read_csv("../../data/experiment2/adults/exp2_adults.csv")

```

#### Wrangle
```{r}
df.exp2_adults = df.exp2_adults %>% 
  mutate(participant = 1:n()) %>% 
  relocate(participant) %>% 
  select(-workerid, -error) %>% 
   pivot_longer(cols = c("video_1", "video_2", "video_3", "video_4", "video_5", "video_6"
   ),
   names_to = "video_version",
   values_to = "video"
  ) %>% 
  pivot_longer(cols = c("video_1_correct_answer", "video_2_correct_answer", "video_3_correct_answer", "video_4_correct_answer", "video_5_correct_answer", "video_6_correct_answer"),
               names_to = "video_version_correct_answer",
               values_to = "correct_answer") %>% 
  pivot_longer(cols = c("video_1_response", "video_2_response", "video_3_response", "video_4_response", "video_5_response", "video_6_response"),
                names_to = "video_version_response",
   values_to = "response") %>% 
  mutate(video_version_correct_answer = str_remove_all(video_version_correct_answer, "_correct_answer"),
         video_version_response = str_remove_all(video_version_response, "_response")) %>% 
  filter(video_version == video_version_correct_answer & video_version_correct_answer == video_version_response) %>% 
  mutate(age = "Adult") %>%
  select(age, participant, scenario = video, correct_answer, response) %>% 
  mutate(rating = if_else(correct_answer == response, 1, 0)) %>%
  rename(trial_response = response,
         response = rating) 

  
```

#### Demographics
```{r message=FALSE, warning=FALSE}

df.exp2_adults_demographics = read_csv("../../data/experiment2/adults/exp2_adults_demographics.csv") 
 

print_demographics_adults(df.exp2_adults_demographics)

```

### STATS
#### Bayesian Regression Models
##### Hypothesis 1 
```{r}

# Overall adults will thank the individual who prevented the worse outcome from occurring at rates that exceed chance

fit.brm3.adults = brm(formula = response ~ 1 + (1 | participant) + (1 | scenario),
     data = df.exp2_adults ,
     family = "bernoulli",
     cores = 4,
     file = "cache/brm3.adults") 

fit.brm3.adults %>% 
  tidy(effects = c("fixed"),
       conf.int = TRUE,
       conf.level = 0.95,
       conf.method = "HPDinterval") %>% 
  mutate(across(.cols = c(estimate, conf.low, conf.high),
                .fns = ~ inv.logit(.))) %>% 
  select(estimate, contains("conf")) 

```

## PLOTS
### Adults + children

```{r fig.height=4, fig.width=6.5, message=FALSE, warning=FALSE}

selected_indices = seq(from = 1, to = 720, by = 6)  
df.exp2_children_selected = df.exp2_children[selected_indices, ]

df.exp2.children.individual = df.exp2_children_selected %>%
    group_by(participant, age_group, age) %>% 
    summarize(pct_correct = mean(total_score) / 6,
              age_mean = mean(age)) %>% 
    mutate(age = as.numeric(age)) %>%
    ungroup() 


df.exp2.adults.individual = df.exp2_adults %>%
  rename(trial_name = scenario) %>%
  group_by(participant) %>%
  summarize(
    total_score = sum(response),
    pct_correct = mean(total_score) / 6) %>%
  ungroup() %>%
  mutate(age_group = "adults",
         age_mean = "8",
         age = "8") %>%
  select(-total_score)


df.exp2.combined.indivudual = rbind(df.exp2.children.individual, df.exp2.adults.individual)


df.exp2.children.means = df.exp2_children_selected %>%
  mutate(pct_correct = total_score / 6) %>% 
  group_by(age_group) %>%
  summarize(
    n = n(),
    age_mean = mean(age),
    response = Hmisc::smean.cl.boot(pct_correct)) %>%
    mutate(index = c("mean", "low", "high")) %>%
    ungroup() %>% 
    pivot_wider(names_from = index,
                values_from = response) 


df.exp2.adults.means = df.exp2.adults.individual %>%
  group_by(age_group) %>%
  summarize(
    n = n(),
    response = Hmisc::smean.cl.boot(pct_correct)) %>%
    mutate(index = c("mean", "low", "high")) %>%
    ungroup() %>% 
    pivot_wider(names_from = index,
                values_from = response) %>%
    mutate(age_mean = as.numeric(8))




df.exp2.combined.means = rbind(df.exp2.children.means, df.exp2.adults.means)


df.exp2.combined.text = df.exp2.combinbed.means %>% 
    select(age_group, age_mean, n) %>% 
    mutate(label = n,
           label = ifelse(age_group == "3", str_c("n = ", n), n),
           y = 0.95)


ggplot() + 
  geom_hline(yintercept = seq(0, 1, 0.25),
             linetype = "dotted",
             color = "black",
             alpha = 0.8) +
  geom_point(data = df.exp2.children.individual,
             mapping = aes(x = age,
                           y = pct_correct,
                           color = age),
             size = 1.5,
             alpha = 1) +
  geom_pointrange(data = df.exp2.combined.means %>% filter(age_group != "adults"),
                  mapping = aes(x = age_mean,
                                y = mean,
                                ymin = low,
                                ymax = high,
                                fill = age_mean),
                  shape = 21, 
                  color = "black", 
                  size = .7,
                  show.legend = FALSE,
                  alpha = 0.8) +
  geom_pointrange(data = df.exp2.combined.means %>% filter(age_group == "adults"),
                  mapping = aes(x = age_mean,
                                y = mean,
                                ymin = low,
                                ymax = high),
                  shape = 23, 
                  fill = "grey",
                  color = "black",
                  size = .7,
                  show.legend = FALSE,
                  alpha = 0.8) +
  geom_text(data = df.exp2.combined.text,
            mapping = aes(x = age_mean,
                          y = y,
                          label = label),
            size = 5,
            color = "black",
            vjust = -1.7) +  
  ggtitle("Who should receive the thank-you sticker?") +
  labs(x = "Age (in years)",
       y = "Probability of thanking the person\n who prevented a negative outcome") +
  coord_cartesian(ylim = c(0, 1)) +
  scale_y_continuous(breaks = seq(0, 1, 0.25),
                     labels = str_c(seq(0, 1, 0.25) * 100, "%"),
                     expand = expansion(add = c(0, 0.12))) + 
  scale_x_continuous(breaks = c(seq(3, 7, 1), 8),
                     labels = c(seq(3, 7, 1), "adult"), 
                     expand = expansion(add = c(0.2, 0.5))) +
  scale_fill_gradientn(colors = children_palette) + 
  scale_color_gradientn(colors = children_palette) + 
  theme(
    axis.text.x = element_text(size = 11),
    axis.text.y = element_text(size = 11),
    axis.title = element_text(size = 11),
    plot.title = element_text(size = 12,
                              hjust = 0.5,
                              face = "bold",
                              margin = margin(b = 15)),
    legend.position = "none",
    legend.text = element_text(size = 11),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    strip.text.x = element_text(margin(b = 0, t = 7, l = 2, r = 3)),
    plot.margin = margin(t = 0.5,
                         l = 0.5,
                         r = 0.5,
                         b = 0.5,
                         unit = "cm"))


ggsave("../../figures/experiment2/exp2_results.pdf",  width = 6.5, height = 4)


```

# EXPERIMENT 3
## CHILDREN
### DATA
#### Read in Data

```{r message=FALSE, warning=FALSE}
df.exp3.children = read_csv("../../data/experiment3/children/exp3_children.csv")%>%
  filter(row_number() >= 0 & row_number() <= 240) %>% 
  select(age, gender, lan_spoken, trial_1, trial_1_score, trial_2, trial_2_score, trial_3, trial_3_score, trial_4, trial_4_score, trial_5, trial_5_score, trial_6, trial_6_score, total_score) %>%
  mutate(age_group = cut(age, breaks = c(3, 4, 5, 6, 7, 8, 9, 10, 11.1),
                         labels = c("3", "4", "5", "6", "7", "8", "9", "10"),
                         right = TRUE)) %>%
  na.omit() %>%
  mutate(participant = 1:n(),
         total_score = as.numeric(total_score),
        across("trial_1"|"trial_2"|"trial_3"|"trial_4"|"trial_5"|"trial_6", ~case_when(
    grepl("exp7_bulb_laundry_correct_(left|right)_(jessie|caleb)", .) ~ "lightbulb",
    grepl("exp7_apple_trashcan_correct_(left|right)_(lily|alex)", .) ~ "apple",
    grepl("exp7_basketball_egg_correct_(left|right)_(charlie|noah)", .) ~ "basketball",
    grepl("exp7_glass_beanbag_correct_(left|right)_(emily|lucas)", .) ~ "glass",
    grepl("exp7_rock_wine_correct_(left|right)_(amber|billy)", .) ~ "rock",
    grepl("exp7_bear_water_correct_(left|right)_(vicky|connor)", .) ~ "bear",
    TRUE ~ .
  ))) %>%
  mutate(across(starts_with("trial"), as.character)) %>%
  pivot_longer(cols = c("trial_1", "trial_2", "trial_3", "trial_4", "trial_5", "trial_6"),
               names_to = "trial_order",
               values_to = "trial_name") %>%
  group_by(participant) %>%
  mutate(trial_score = case_when(
    trial_order == "trial_1" ~ trial_1_score,
    trial_order == "trial_2" ~ trial_2_score,
    trial_order == "trial_3" ~ trial_3_score,
    trial_order == "trial_4" ~ trial_4_score,
    trial_order == "trial_5" ~ trial_5_score,
    trial_order == "trial_6" ~ trial_6_score,
  )) %>%
  mutate(trial_score = as.numeric(trial_score)) %>%
  mutate(trial_order = ifelse(trial_order == "trial_1", "trial 1",
                               ifelse(trial_order == "trial_2", "trial 2",
                                      ifelse(trial_order == "trial_3", "trial 3",
                                             ifelse(trial_order == "trial_4", "trial 4",
                                                    ifelse(trial_order == "trial_5", "trial 5",
                                                           ifelse(trial_order == "trial_6", "trial 6", trial_order))))))) %>%
  mutate(side = case_when(
    trial_name %in% c("basketball", "bear", "apple", "rock") ~ "side with things",
    trial_name %in% c("lightbulb", "glass") ~ "side with nothing",
    TRUE ~ "Uncategorized" 
  )) %>%
  mutate(prefer_things = case_when(
    side == "side with things" & trial_score == 1 ~ 1,
    side == "side with nothing" & trial_score == 0 ~ 1,
    TRUE ~ 0
  )) %>%
  select(participant, gender, lan_spoken, age, age_group, trial_order, trial_name, trial_score, side, prefer_things, total_score)

```

## Demographics
```{r message=FALSE, warning=FALSE}

print_demographics(df.exp3.children)

```
# STATS
## Bayesian Regression Models
### Hypothesis 1 
```{r message=FALSE, warning=FALSE}

# Overall children will thank the individual who prevented the worse outcome from occurring at rates that exceed chance

fit.brm5.children = brm(formula = trial_score ~ 1 + (1 | participant) + (1 | trial_name),
     data = df.exp3.children,
     family = "bernoulli", 
     cores = 4,
     file = "cache/brm5_children") 

fit.brm5.children %>% 
  tidy(effects = c("fixed"),
       conf.int = TRUE,
       conf.level = 0.95,
       conf.method = "HPDinterval") %>% 
  mutate(across(.cols = c(estimate, conf.low, conf.high),
                .fns = ~ inv.logit(.))) %>% 
  select(estimate, contains("conf"))

```

### Hypothesis 2
```{r message=FALSE, warning=FALSE}

# The tendency to thank the individual who prevented the worse potential outcome from occurring will increase with age

fit.brm5.children.age = brm(formula = trial_score ~ 1 + age + (1 | participant) + (1 | trial_name),
     data = df.exp3.children,
     family = "bernoulli", 
     cores = 4,
     file = "cache/brm5_children_age") 

fit.brm5.children.age %>% 
  tidy(effects = c("fixed"),
       conf.int = TRUE,
       conf.level = 0.95,
       conf.method = "HPDinterval") %>% 
  filter(term == "age") %>%
  select(estimate, contains("conf")) 

```

## ADULTS
### DATA
#### Read in Data

```{r message=FALSE, warning=FALSE}


setwd("../../data/experiment3/adults")

file_names = list.files(pattern="*.csv")
file_names = file_names[!file_names %in% c("m6f0atxjkr_same_participant_as_lhjb5uy11r.csv")]
all_data = do.call(rbind, lapply(file_names, read.csv)) %>%
  distinct()

write.csv(all_data, "exp3_adult.csv", row.names = FALSE)

df.exp3.adults = read_csv("exp3_adult.csv")

```

#### Wrangle
```{r}
df.exp3.adults = df.exp3.adults %>% 
  mutate(participant = 1:n()) %>% 
  relocate(participant) %>% 
   pivot_longer(cols = c("video_1", "video_2", "video_3", "video_4", "video_5", "video_6"
   ),
   names_to = "video_version",
   values_to = "video"
  ) %>% 
  pivot_longer(cols = c("video_1_correct_answer", "video_2_correct_answer", "video_3_correct_answer", "video_4_correct_answer", "video_5_correct_answer", "video_6_correct_answer"),
               names_to = "video_version_correct_answer",
               values_to = "correct_answer") %>% 
  pivot_longer(cols = c("video_1_response", "video_2_response", "video_3_response", "video_4_response", "video_5_response", "video_6_response"),
                names_to = "video_version_response",
   values_to = "response") %>% 
  mutate(video_version_correct_answer = str_remove_all(video_version_correct_answer, "_correct_answer"),
         video_version_response = str_remove_all(video_version_response, "_response")) %>% 
  filter(video_version == video_version_correct_answer & video_version_correct_answer == video_version_response) %>% 
  mutate(age = "Adult") %>%
  select(age, participant, scenario = video, correct_answer, response, feedback) %>% 
  mutate(rating = if_else(correct_answer == response, 1, 0)) %>%
  rename(trial_response = response,
         response = rating) 

```


#### Demographics
```{r message=FALSE, warning=FALSE}

# read in data
df.exp3.adults_demographics = read_csv("../../data/experiment3/adults/exp3_adult.csv")

# print demographics
print_demographics_adults(df.exp3.adults_demographics)

```

### STATS
#### Bayesian Regression Models
##### Hypothesis 1 
```{r message=FALSE, warning=FALSE}

# Overall adults will thank the individual who prevented the worse outcome from occurring at rates that exceed chance

fit.brm5.adults = brm(formula = response ~ 1 + (1 | participant) + (1 | scenario),
     data = df.exp3.adults,
     family = "bernoulli",
     cores = 4,
     file = "cache/brm5.adults") 

fit.brm5.adults %>% 
  tidy(effects = c("fixed"),
       conf.int = TRUE,
       conf.level = 0.95,
       conf.method = "HPDinterval") %>% 
  mutate(across(.cols = c(estimate, conf.low, conf.high),
                .fns = ~ inv.logit(.))) %>% 
  select(estimate, contains("conf")) 

```
## PLOTS
### Adults + children

```{r fig.height=4, fig.width=8.5, message=FALSE, warning=FALSE}

selected_indices = seq(from = 1, to = 1440, by = 6)  
df.exp3_children_selected = df.exp3.children[selected_indices, ]

df.exp3.children.individual = df.exp3_children_selected %>%
    group_by(participant, age_group, age) %>% 
    summarize(pct_correct = mean(total_score) / 6,
              age_mean = mean(age)) %>% 
    mutate(age = as.numeric(age)) %>%
    ungroup() 

df.exp3.adults.individual = df.exp3.adults %>%
  rename(trial_name = scenario) %>%
  group_by(participant) %>%
  summarize(
    total_score = sum(response),
    pct_correct = mean(total_score) / 6) %>%
  ungroup() %>%
  mutate(age_group = "adults",
         age_mean = "12",
         age = "12") %>%
  select(-total_score)




df.exp3.combined.individual = rbind(df.exp3.children.individual, df.exp3.adults.individual)


df.exp3.children.means = df.exp3_children_selected %>%
  mutate(pct_correct = total_score / 6) %>% 
  group_by(age_group) %>%
  summarize(
    n = n(),
    age_mean = mean(age),
    response = Hmisc::smean.cl.boot(pct_correct)) %>%
    mutate(index = c("mean", "low", "high")) %>%
    ungroup() %>% 
    pivot_wider(names_from = index,
                values_from = response) 


df.exp3.adults.means = df.exp3.adults.individual %>%
  group_by(age_group) %>%
  summarize(
    n = n(),
    response = Hmisc::smean.cl.boot(pct_correct)) %>%
    mutate(index = c("mean", "low", "high")) %>%
    ungroup() %>% 
    pivot_wider(names_from = index,
                values_from = response) %>%
    mutate(age_mean = as.numeric(12))




df.exp3.combined.means = rbind(df.exp3.children.means, df.exp3.adults.means)


df.exp3.combined.text = df.exp3.combined.means %>% 
    select(age_group, age_mean, n) %>% 
    mutate(label = n,
           label = ifelse(age_group == "3", str_c("n = ", n), n),
           y = 0.95)

ggplot() + 
  geom_hline(yintercept = seq(0, 1, 0.25),
             linetype = "dotted",
             color = "black",
             alpha = 0.8) +
  geom_point(data = df.exp3.children.individual,
             mapping = aes(x = age,
                           y = pct_correct,
                           color = age),
             size = 1.5,
             alpha = 1) +
  geom_pointrange(data = df.exp3.combined.means %>% filter(age_group != "adults"),
                  mapping = aes(x = age_mean,
                                y = mean,
                                ymin = low,
                                ymax = high,
                                fill = age_mean),
                  shape = 21, 
                  color = "black", 
                  size = .7,
                  show.legend = FALSE,
                  alpha = 0.8) +
  geom_pointrange(data = df.exp3.combined.means %>% filter(age_group == "adults"),
                  mapping = aes(x = age_mean,
                                y = mean,
                                ymin = low,
                                ymax = high),
                  shape = 23, 
                  fill = "grey",
                  color = "black",
                  size = .7,
                  show.legend = FALSE,
                  alpha = 0.8) +
  geom_text(data = df.exp3.combined.text,
            mapping = aes(x = age_mean,
                          y = y,
                          label = label),
            size = 5,
            color = "black",
            vjust = -1.7) +  
  ggtitle("Who should receive the thank-you sticker?") +
  labs(x = "Age (in years)",
       y = "Probability of thanking the person\n who prevented a negative outcome") +
  coord_cartesian(ylim = c(0, 1)) +
  scale_y_continuous(breaks = seq(0, 1, 0.25),
                     labels = str_c(seq(0, 1, 0.25) * 100, "%"),
                     expand = expansion(add = c(0, 0.12))) + 
  scale_x_continuous(breaks = c(seq(3, 11, 1), 12),
                     labels = c(seq(3, 11, 1), "adult"), 
                     expand = expansion(add = c(0.2, 0.5))) +
  scale_fill_gradientn(colors = children_palette) + 
  scale_color_gradientn(colors = children_palette) + 
  theme(
    axis.text.x = element_text(size = 11),
    axis.text.y = element_text(size = 11),
    axis.title = element_text(size = 11),
    plot.title = element_text(size = 12,
                              hjust = 0.5,
                              face = "bold",
                              margin = margin(b = 15)),
    legend.position = "none",
    legend.text = element_text(size = 11),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    strip.text.x = element_text(margin(b = 0, t = 7, l = 2, r = 3)),
    plot.margin = margin(t = 0.5,
                         l = 0.5,
                         r = 0.5,
                         b = 0.5,
                         unit = "cm"))



ggsave("../../figures/experiment3/exp3_results.pdf",  width = 8.5, height = 4)



```
# SESSION INFO

```{r}
sessionInfo()
```